From physical principles to ???

### 2016-01-14 2015CompleteDerivation/completeDerivation.tex
Adding case with no env. Intro to state var

We start by fixing a \emph{physical system} we want to study, that is something we can interact with and perform measurements on (e.g. a cat, a planet, water). We call \emph{environment} everything else. We set what particular aspect we want to study (e.g. the trajectory when falling, the motion around a star, its flow in a pipe). We call \emph{state} a particular configuration in time of the portion under study. Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word. With this in mind, we introduce the following:

@@

As the terms used have different meaning in different contexts, let's first clarify them. By deterministic (and reversible) we mean that future (and past) states are uniquely identified by the present state. We can then think of the state of the system as the part that interacts deterministically and reversibly with the environment (and itself) while the unstated part as interacting non-deterministically with the environment (and itself). We call this setup the \emph{fundamental model of physics}, as it is the simplest model that still captures the crucial aspects of studying and writing laws for a physical system.

@@

@@

>>>>

We start by fixing a \emph{physical system} we want to study, that is something we can interact with and perform measurements on (e.g. a planet, water). We call \emph{environment} everything else. We set what particular aspect we want to study (e.g. the motion around a star, the flow in a pipe). We call \emph{state} a physically distinguishable configuration of the portion under study at a particular time. Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word. Note that the environment plays an essential role here as it's what allows us to define two states as physically distinguishable: we can find an external process (i.e. a different environment) whose outcome changes depending on the different state. As such processes are what we can use to perform measurements, we consider the experimental apparatus (and us performing measurements) independent of the system, part of the environment. This is true even in the case of general relativity, where we can imagine multiple researchers on small spaceships collecting data without greatly influencing the motion of stars and planets. We call this triad of state, unstated part and environment the \emph{fundamental model of physics}, as it is the simplest model that still captures the crucial aspects of studying and writing laws for a physical system.\footnote{The case where the physical system is the whole universe and there is no environment presents practical problems (e.g. What physical device can we use to store and process the state of the whole universe to make predictions and compare?) and conceptual challenges (e.g. How do we define physically distinguishable? Can the physical laws predict which measurements are we going to make? Does that limit what is actually distinguishable?) that the current physical theories do not seem to be equipped to address, and therefore will be absent from our discussion.} With this in mind, we introduce the following:

@@

As the terms used have different meaning in different contexts, let's first clarify them. By deterministic (and reversible) we mean that future (and past) states are uniquely identified by the present state. We can then think of the state of the system as the part that interacts deterministically and reversibly with the environment (and itself) while the unstated part as interacting non-deterministically with the environment (and itself).

@@

To identify and name states, we will typically use a set of quantities, typically numbers. For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$. We call each of these quantities \emph{state variables} and \emph{state vector} a set of state variables that can identify one and only one state. We avoid the term coordinates, even though that's what they are mathematically, as it will create confusion with spatial coordinates. We also avoid the term observable or measurable, as not all state variables map directly to the outcome of an experiment (e.g. conjugate momentum as it is not gauge invariant). It's clear, though, that even in those cases there must exist, at least in principle, a way to distinguish between the possible values of the state variable (e.g once we fix the vector potential, conjugate momentum can be measured via the kinetic momentum. Therefore we can assume that all state variables are in the end physically distinguishable, yet we do not assume all state variables have equal standing in physical meaning. Their purpose is just to label states, and the chosen name is a good reminder.

@@

\begin{prop}\label{state_vector}
TODO: State vector always exists
\end{prop}

When assembling multiple state variable into a state vector, it is important to understand how they relate to each other. Consider the orbital of an electron in a hydrogen item, which is identified by the quantum numbers $n$, $l$, $m$ and $s$. For each combination of $n$, $l$ and $m$, the spin $s$ can have to value. The choice of spin is independent from the rest. The choice of $l$, though, depends on the choice of $n$: for $n=1$ only $l=s$ is available; for $n=2$ we can choose $l=s$ or $l=p$. The choices are not independent. That is: two or more variables are independent if there always exist a state for any possible combination.
